### 2021-12-06 input/proposal.tex
reworking theo 1

Unfortunately, \processes do not share a common global knowledge of
the network state. For \processes to eventually ($\eventually$) reach
consistent partitioning, each source $s$ must send notifications
$\alpha_s$ to all \processes that are closer to it than any other
source, along with enough control information to allow them to decide
of their closest source.

\begin{theorem}[\label{theo:fb}\underline{F}orwarding of \underline{B}est (FB)
    $\implies\eventually \textup{CP}$]
    % 
    Assuming reliable communication links ($s_{xy}(m) \iff \eventually
    r_{yx}(m)$), and a total order on messages ($<$), \processes
    eventually reach consistent partitioning if each \process delivers
    its best message among \underline{r}eceived messages $R_x$, and
    forwards its best message among \underline{d}elivered messages
    $D_x$.
    % , \ie
    % the message $\alpha_s^d$ with the smallest distance $d$ to the
    % source $s$
    % ($(\alpha_s^d = \min R_x) \implies (d_x(\alpha_s^d) \wedge
    % f_x(\TODO{\alpha_s^d}))$).
\end{theorem}

\begin{proof}
  % When there are no sources, no \process receives, nor delivers hence
  % forwards any message. \Processes do not belong to any partition.

  \noindent \TODO{use total order ($<$) and sets $R_p D_p$} Whenever a
  \process $s$ becomes a source, it delivers its own message
  $d_s(\alpha_{s})$. Whatever its set of received messages, its own
  partition $P_s$ becomes and remains its closest partition
  $\alpha_{s}^{|\Pi_{ss}|} = \min D_s$ since $\forall x \neq s:
  |\Pi_{s s}| < |\Pi_{s x}|$.
  %
  \noindent Such a source $s$ forwards $\alpha_{s}$ to its neighbors.
  Since communication links are reliable, neighboring \processes
  eventually receive such a notification
  $\forall \langle s, y \rangle \in E \iff \eventually
  r_y(\alpha_{s}^{d})$. Most importantly, whatever the order of
  received messages, every \process $q$ in this neighborhood such that
  there exists no closer source $s'$ than the received one $s$,
  delivers it, since $w_{sq} < |\Pi_{s'q}|$.
  %
  \noindent They forward it and their respective neighbors eventually
  receive it. By transitivity, the message originating from $s$
  reaches all \processes belonging to $P_s$ at least through their
  shortest paths: $\smash{\forall q' \in V, s, s' \in S: |\Pi_{s q'}| <
  |\Pi_{s' q'}| \implies \eventually d_{q'}(\alpha_{s}^{|\Pi_{sq'}|})}$.
  Since there exists only one best sum of weights per \process that
  can never be retracted, \processes eventually reach consistent
  partitioning.
  %% /!\ not equivalence for there exists other implementations.
  %% \item [$CP \implies BEF$:] By contradiction, if a \process $q \in
  %%   \Pi_{sp} = [s, \ldots, q, q', \ldots, p]$ with $s, q, p \in P_s$
  %%   does not forward its received $\alpha_s^{|\Pi_{sq}|}$, then
  %%   following \processes from $q'$ to $p$ may mistake another
  %%   partition for their  because it needs the weight $W_{pq}$.
  %% \end{asparadesc}
\end{proof}

@@

Algorithm~\ref{algo:add} shows the instructions that implement
forwarding of best to reach consistent partitioning when
\begin{inparaenum}[(i)]
\item weights are scalar values piggybacked in messages and
  accumulated over forwarding,
\item \processes only add new partitions to the system,
\item and \processes never join, crash nor leave the system.
\end{inparaenum}
For the sake of simplicity and to remove obvious loops, a \process
does not forward a notification to its sender
(Line~\ref{line:accumulator}).  Figure~\ref{fig:add} illustrates the
behavior of this algorithm on a system comprising 4 \processes $a$,
$b$, $c$, and $d$. \Process~$a$ and \Process~$d$ become the sources of
their partition.  They \NAMEB a notification \TODO{\underline{a}dd
  message}: $\alpha_a^0$ and $\alpha_d^0$. They initialize their own
state with the lowest possible bound $0$ (see
Line~\ref{line:lowestbound}), and send a message to each of their
neighbors by accumulating the corresponding edge weight (see
Line~\ref{line:accumulator}). In Figure~\ref{fig:addC}, \Process~$b$
receives $\alpha_{d}^{1}$. Since it improves its own partition
distance, it keeps it and forwards it to its neighbors. In
Figure~\ref{fig:addD}, $b$ discards $\alpha_{a}^{2}$, for it does not
improve its partition distance. \Processes $c$ and $d$ will never
acknowledge that Source~$a$ exists. Ultimately, \processes discard
last transiting messages. Despite the obvious lack of traffic
optimization, the system reaches consistent partitioning.

>>>>

Unfortunately, \processes do not share a common global knowledge of
the network state. For \processes to eventually ($\eventually$) reach
consistent partitioning, each source $s$ must send notifications
$\alpha_s$ to all \processes that are closer to it than any other
source, along with enough metadata to allow them to decide of their
closest source. A simple implementation consists in ensuring that each
\process $x$ eventually receives $\alpha_s^{|\Pi_{sx}|}$ where
$\Pi_{sx}$ is the \emph{best} path from any source to it; the value
$|\Pi_{sx}|$ being built over forwarding along this path.

\begin{theorem}[\label{theo:fb}\underline{F}orwarding of \underline{B}est (FB)
    $\implies\eventually \textup{CP}$]
    % 
    Assuming that each \process $x$ stores its outgoing weights
    ($\forall \langle x, y \rangle \in E: \pi_{xy} \in \sigma_x$), a
    total order on messages based on weights ($m^d_s < m^{d'}_{s'}$
    when $d < d' \vee (d = d' \wedge s< s')$), reliable communication
    links ($s_{xy}(m) \iff r_{yx}(m)$), \processes eventually reach
    consistent partitioning if each \process delivers its best (\ie
    smallest) message among \underline{r}eceived messages $R_x$, and
    forwards its best message among \underline{d}elivered messages
    $D_x$ such that $f_x(\alpha^d_s) \implies \forall \langle x, y
    \rangle \in E: s_{xy}(\alpha_s^d \oplus ^{w_{xy}})$, \ie by
    accumulating the respective weight of used edges.
    % ($(\alpha_s^d = \min R_x) \implies (d_x(\alpha_s^d) \wedge
    % f_x(\TODO{\alpha_s^d}))$).
\end{theorem}

\begin{proof}
  % When there are no sources, no \process receives, nor delivers hence
  % forwards any message. \Processes do not belong to any partition.
  \noindent \TODO{use total order ($<$) and sets $R_p D_p$} Whenever a
  \process $s$ becomes a source, it delivers its own message
  $d_s(\alpha_{s}^{|\pi_{ss}|})$. Whatever its set of received
  messages, its own partition $P_s$ becomes and remains its closest
  partition $\alpha_{s}^{|\Pi_{ss}|} = \min D_s$ since $\forall x \neq
  s: |\Pi_{s s}| < |\Pi_{s x}|$.
  %
  \noindent Such a source $s$ forwards $\alpha_{s}$ to its neighbors.
  Since communication links are reliable, neighboring \processes
  eventually receive such a notification
  $\forall \langle s, y \rangle \in E \iff \eventually
  r_y(\alpha_{s}^{d})$. Most importantly, whatever the order of
  received messages, every \process $q$ in this neighborhood such that
  there exists no closer source $s'$ than the received one $s$,
  delivers it, since $w_{sq} < |\Pi_{s'q}|$.
  %
  \noindent They forward it and their respective neighbors eventually
  receive it. By transitivity, the message originating from $s$
  reaches all \processes belonging to $P_s$ at least through their
  shortest paths: $\smash{\forall q' \in V, s, s' \in S: |\Pi_{s q'}| <
  |\Pi_{s' q'}| \implies \eventually d_{q'}(\alpha_{s}^{|\Pi_{sq'}|})}$.
  Since there exists only one best sum of weights per \process that
  can never be retracted, \processes eventually reach consistent
  partitioning.
  %% /!\ not equivalence for there exists other implementations.
  %% \item [$CP \implies BEF$:] By contradiction, if a \process $q \in
  %%   \Pi_{sp} = [s, \ldots, q, q', \ldots, p]$ with $s, q, p \in P_s$
  %%   does not forward its received $\alpha_s^{|\Pi_{sq}|}$, then
  %%   following \processes from $q'$ to $p$ may mistake another
  %%   partition for their  because it needs the weight $W_{pq}$.
  %% \end{asparadesc}
\end{proof}

@@

Algorithm~\ref{algo:add} shows the instructions that implement a
\emph{reactive} forwarding of best to reach consistent partitioning
when
\begin{inparaenum}[(i)]
\item weights are scalar values piggybacked in messages and
  accumulated over forwarding,
\item \processes only add new partitions to the system,
\item and \processes never join, crash nor leave the system.
\end{inparaenum}
\TODO{We only keep last.}  For the sake of simplicity and to remove
obvious loops, a \process does not forward a notification to its
sender (Line~\ref{line:accumulator}).  Figure~\ref{fig:add}
illustrates the behavior of this algorithm on a system comprising 4
\processes $a$, $b$, $c$, and $d$. \Process~$a$ and \Process~$d$
become the sources of their partition.  They \NAMEB a notification
\TODO{\underline{a}dd message}: $\alpha_a^0$ and $\alpha_d^0$. They
initialize their own state with the lowest possible bound $0$ (see
Line~\ref{line:lowestbound}), and send a message to each of their
neighbors by accumulating the corresponding edge weight (see
Line~\ref{line:accumulator}). In Figure~\ref{fig:addC}, \Process~$b$
receives $\alpha_{d}^{1}$. Since it improves its own partition
distance, it keeps it and forwards it to its neighbors. In
Figure~\ref{fig:addD}, $b$ discards $\alpha_{a}^{2}$, for it does not
improve its partition distance. \Processes $c$ and $d$ will never
acknowledge that Source~$a$ exists. Ultimately, \processes discard
last transiting messages. Despite the obvious lack of traffic
optimization, the system reaches consistent partitioning.
